---
title: "Climate sensitivity and management of endangered animals"
output: 
  flexdashboard::flex_dashboard:
    css: custom.css
    social: menu
    df_print: paged
    navbar:
    - {title: "CCI", align: right, href: "https://defenders-cci.org"}
    - {title: "", icon: "fa-question-circle fa-lg", align: right, href: "mailto:adelach@defenders.org?subject=ESA climate sensitivity"}
    - {title: "", icon: "fa-github fa-lg", align: right, href: "https://github.com/Defenders-CCI/"}
runtime: shiny
---

```{r setup, include=FALSE}
library(dplyr)
library(flexdashboard)
library(ggplot2)
library(plotly)
library(rio)
library(shiny)
library(shinydashboard)
library(stringr)
library(viridis)

dat <- readRDS("data/sensitivity_data_prepped_v2.rds")
dis <- readRDS("data/discussion_data_prepped_v2.rds")

sens_dwnld <- readRDS("data/Delach_et_al_2019_sensitivity_data.rds")
disc_dwnld <- readRDS("data/Delach_et_al_2019_discussion_data.rds")
sensmeta <- readRDS("data/sensitivity_metadata.rds")
dismeta <- readRDS("data/discussion_metadata.rds")

taxa <- c("All", unique(dat$Taxon) %>% sort())
subtaxa <- c("All", unique(dat$Subtaxon) %>% sort())
region <- c("All", unique(dat$Lead_Region) %>% sort())
state <- c("All", 
           dis$State %>% 
             str_split(", | |,") %>% 
             unlist() %>% 
             str_trim() %>% 
             unique() %>%
             sort())
species <- unique(dat$species) %>% sort()
```

Overview
=====================================================


Filters {.sidebar}
-----------------------------------------------------
### Biodiversity

```{r}
selectInput(
  "taxa_1",
  "Vertebrate?",
  taxa
)

selectInput(
  "taxa_2",
  "Species group",
  subtaxa
)
```

### Area

```{r}
selectInput(
  "region_1",
  "Agency/region",
  region
)

selectInput(
  "region_2",
  "State/territory",
  state
)
```

```{r}
usr_sel <- function(df, tx1, tx2, re1, re2) {
  if(tx1 != "All") {
    df <- filter(df, Taxon == tx1)
  }
  if(tx2 != "All") {
    df <- filter(df, Subtaxon == tx2)
  }
  if(re1 != "All") {
    df <- filter(df, Lead_Region == re1)
  }
  if(re2 != "All") {
    df <- filter(df, grepl(pattern = re2, State))
  }
  if(dim(df)[1] == 0) stop("No data match filters; please update selections.")
  return(df)
}

dat_select <- reactive({
  cur_dat <- usr_sel(dat, input$taxa_1, input$taxa_2,
                     input$region_1, input$region_2)
})

dis_select <- reactive({
  cur_dat <- usr_sel(dis, input$taxa_1, input$taxa_2,
                     input$region_1, input$region_2)
})

```


```{r}
helpText("Use the drop-downs to filter the data to your species group
         or area of interest. A warning will show if the combination of 
         filters gives no results.")
```

### Background
```{r}
helpText("This study examines the sensitivity of U.S. endangered animals 
         to climate change and at what level agency plans discuss climate 
         change threats and management. See ",a(href="https://defenders-cci.org/app/ESA_climate/#section-data", target="_blank", "Data+")," for details.")
```

Col 1
-----------------------------------------------------

### Degree of climate sensitivity

```{r}
# renderPlotly({
#   o1 <- dat %>% distinct(Scientific_Name, .keep_all = TRUE)
#   d1 <- dat_select() %>% distinct(Scientific_Name, .keep_all = TRUE)
#   p <- ggplot(data = d1, aes(n_sens_factors)) +
#     geom_bar(data = o1, aes(n_sens_factors), alpha = 0.3) + 
#     geom_bar(alpha = 0.7, fill = viridis(1)) + 
#     labs(x = "# Sensitivity Factors",
#          y = "# Species") +
#     scale_fill_viridis(discrete = TRUE) +
#     theme_bw() +
#     theme(legend.position = "none")
#   ggplotly(p) %>%
#     layout(legend = list(orientation = "h"))
# })
renderPlotly({
  o1 <- dat %>% distinct(Scientific_Name, .keep_all = TRUE)
  d1 <- dat_select() %>% distinct(Scientific_Name, .keep_all = TRUE)
  plot_ly(
    data = o1,
    type = "histogram",
    x = o1$n_sens_factors,
    marker = list(color = "rgba(0,0,0,0.2)"),
    name = "All data"
  ) %>% 
    add_histogram(x = d1$n_sens_factors,
                  marker = list( color = "#C41230" ),
                  name = "Selected data") %>%
    layout(barmode = "overlay",
           xaxis = list(title = "# Sensitivity Factors"),
           yaxis = list(title = "# Species"))
})
```

### Ranking of climate sensitivity threats

```{r}
renderPlotly({
  o1 <- dat %>% 
    filter(Sens_Disc == "y")
  nsp <- table(o1$Factor) %>% 
    as.data.frame() %>% 
    arrange(desc(Freq))
  sort_levels <- nsp$Var1[order(nsp$Freq, decreasing = TRUE)]
  nsp$Var1 <- factor(nsp$Var1, levels = sort_levels)
  d1 <- dat_select() %>%
    filter(Sens_Disc == "y")
  # d1 <- filter(dat, Taxon == "Vertebrate") %>%
  #   filter(Sens_Disc == "y")
  fnsp <- table(d1$Factor) %>% 
    as.data.frame()
  fnsp$Var1 <- factor(fnsp$Var1, levels = sort_levels)
  plot_ly(
    data = nsp,
    type = "bar",
    x = ~nsp$Var1,
    y = ~nsp$Freq,
    marker = list(color = "rgba(0,0,0,0.2)")
  ) %>%
    add_bars(
      data = fnsp,
      x = ~fnsp$Var1,
      y = ~fnsp$Freq,
      marker = list( color = "#387C2B" )
    ) %>%
    layout(barmode = "overlay",
           xaxis = list(title = ""),
           yaxis = list(title = "# Species"),
           showlegend = FALSE)
})
```

Col 2
------------------------------------------------------

### Is climate change disussed as a potential threat?

```{r}
renderPlotly({
  o1 <- dis
  n_o1 <- table(o1$CC_Discuss) %>% as.data.frame()
  d1 <- dis_select()
  n_d1 <- table(d1$CC_Discuss) %>% as.data.frame()
  plot_ly(
    data = n_o1,
    type = "bar",
    x = ~n_o1$Var1,
    y = ~n_o1$Freq,
    marker = list(color = "rgba(0,0,0,0.2)")
  ) %>%
    add_bars(
      data = n_d1,
      x = ~n_d1$Var1,
      y = ~n_d1$Freq,
      marker = list( color = "#E36F1E" )
    ) %>%
    layout(barmode = "overlay",
           xaxis = list(title = ""),
           yaxis = list(title = "# Species"),
           showlegend = FALSE)
})
```

### At what level is climate change management discussed?

```{r}
renderPlotly({
  dis1 <- subset(dis, Discuss_Level!="Excluded from analysis") #excluding this category (n=39) 
  dis1$Discuss_Level <- as.factor(dis1$Discuss_Level)
  dis1$Discuss_Level <- droplevels(dis1$Discuss_Level)
  dis1$Discuss_Level <- as.character(dis1$Discuss_Level)
  o2 <- table(dis1$Discuss_Level) %>% as.data.frame() %>%
    arrange(desc(Freq))
  o2$Var1 <- str_replace(o2$Var1, "\n", "<br>")
  d2 <- table(dis_select()$Discuss_Level) %>% as.data.frame()
  d2$Var1 <- str_replace(d2$Var1, "\n", "<br>")
  b1 <- left_join(o2, d2, by = "Var1")
  b1$Var1 <- factor(b1$Var1,levels=c("No discussion","No threat,<br>no action needed","Further study","Action"))
  observe({print(b1)})
  plot_ly(
    data = b1,
    type = "bar",
    x = ~b1$Var1,
    y = ~b1$Freq.x,
    marker = list(color = "rgba(0,0,0,0.2)")
  ) %>%
    add_bars(
      data = b1,
      y = ~b1$Freq.y,
      marker = list( color = "#7A3777" ) #6B4577
    ) %>%
    layout(barmode = "overlay",
           xaxis = list(title = ""),
           yaxis = list(title = "# Species"),
           showlegend = FALSE)
})
```


Species {data-orientation=rows}
=====================================================

Filter {.sidebar}
-----------------------------------------------------

### Species

```{r}
species <- unique(dat$species) %>% sort()

selectInput(
  "species_1",
  label = NULL,
  species
)
```

```{r}
cur_sp <- reactive({
  filter(dat, species == input$species_1)
})

cur_dis <- reactive({
  filter(dis, combo_name == input$species_1)
})
```

Row 0a
-----------------------------------------------------

<h1>`r reactive(input$species_1)`</h1>

Row 1
-----------------------------------------------------

### Taxonomic group {.value-box}

```{r}
# observe({ print(cur_dis()$Subtaxon) })
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = cur_dis()$Subtaxon,
    color = "info",
    icon = case_when(
      cur_dis()$Subtaxon == "Bird" ~ "fa-crow",
      cur_dis()$Subtaxon == "Fish" ~ "fa-fish",
      cur_dis()$Subtaxon == "Arthropod" ~ "fa-spider",
      cur_dis()$Subtaxon == "Amphibian" ~ "fa-frog",
      cur_dis()$Subtaxon == "Reptile" ~ "fa-hand-lizard",
      cur_dis()$Subtaxon == "Mammal" ~ "fa-paw",
      cur_dis()$Subtaxon == "Mollusck" ~ "fa-circle"
    )
  )
})
```

### Lead region {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = cur_dis()$Lead_Region,
    color = "info",
    icon = "fa-map"
  )
})
```

### Year discussed {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = cur_dis()$Date,
    color = "info",
    icon = "fa-calendar"
  )
})
```

Row 1b
-----------------------------------------------------

<h1>Climate Sensitivities</h1>

Row 2
-----------------------------------------------------

### `r reactive({cur_sp()$Factor[1] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[1] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[1] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[1] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

### `r reactive({ cur_sp()$Factor[2] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[2] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[2] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[2] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

### `r reactive({ cur_sp()$Factor[3] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[3] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[3] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[3] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

### `r reactive({ cur_sp()$Factor[4] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[4] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[4] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[4] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

Row 3 
----------------------------------------------------

### `r reactive({cur_sp()$Factor[5] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[5] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[5] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[5] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

### `r reactive({ cur_sp()$Factor[6] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[6] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[6] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[6] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

### `r reactive({ cur_sp()$Factor[7] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[7] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[7] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[7] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

### `r reactive({ cur_sp()$Factor[8] })` {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_sp()$Sens_Disc[8] == "y", "Yes", "No"),
    color = if_else(cur_sp()$Sens_Disc[8] == "y", "danger", "success"),
    icon = if_else(cur_sp()$Sens_Disc[8] == "y", "fa-check-circle", "fa-times-circle")
  )
})
```

Row 2b
-----------------------------------------------------

<h1>Climate Discussion</h1>

Row 4
----------------------------------------------------

### Climate change discussed as potential threat? {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = if_else(cur_dis()$CC_Discuss == "Yes", "Yes", "No"),
    color = if_else(cur_dis()$CC_Discuss == "Yes", "success", "danger"),
    icon = if_else(cur_dis()$CC_Discuss == "Yes", "fa-check-circle", "fa-times-circle")
  )
})
```

### Level of climate change management discussed {.value-box}

```{r}
flexdashboard::renderValueBox({
  flexdashboard::valueBox(
    value = cur_dis()$Discuss_Level,
    color = if_else(
      cur_dis()$Discuss_Level %in% c("Action"), 
      "success", "danger"),
    icon = if_else(
      cur_dis()$Discuss_Level %in% c("Action"), 
      "fa-comment", "fa-times-circle")
  )
})
```



Data+
=====================================================

col 1 {data-width:250}
-----------------------------------------------------

col 2 {data-width=500}
-----------------------------------------------------

<h4>Data files</h4>

<a class='btn btn-default shiny-download-link shiny-bound-output' href='https://osf.io/r9uca/' style="width:250px" target="_blank">Access data files</a>

Data and R script files are archived on Open Science Framework.



<h4>Paper</h4>

<a class='btn btn-default shiny-download-link shiny-bound-output' href='https://www.nature.com/articles/s41558-019-0620-8' style="width:250px" target="_blank">Access paper</a>

Citation: Delach A, Caldas A, Edson K, Krehbiel R, Murray S, Theoharides K, Vorhees L, Malcom JW, Salvo M, Miller JRB. 2019. Agency plans are inadequate to conserve US endangered species under climate change. Nature Climate Change. doi: 10.1038/s41558-019-0620-8.

```{r}
# DT::renderDataTable({
#   # tab1 <- select(dat, Question, Explanation, Sens_Disc)
#   DT::datatable(dat)
# })
```

col 3 {data-width:250}
-----------------------------------------------------
